Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-07 intro.tex
part of introduction: conceptual overview

>>>>

\subsection{Conceptual overview: Potential for the gradient operator and the corresponding Poincar\'e--Friedrichs inequality} \label{subsection:intro_grad}

Let $\Omega \subseteq \bbR^n$, $n \geq 1$, be a bounded connected open set. Given any $1 \leq p \leq \infty$, we let $L^{p}(\Omega)$ denote the Lebesgue space over $\Omega$ with integrability exponent $p$. Then $\bfL^{p}(\Omega) := L^{p}(\Omega)^{n}$ is the corresponding Lebesgue space of vector  fields.

Let $W^{1,p}(\Omega)$ denote the first-order Sobolev space over $\Omega$ with integrability exponent $p$, 
\begin{align} \label{equation:W1p}
    W^{1,p}(\Omega) := \left\{u \in L^{p}(\Omega) \cye{\suchthat} \nabla u \in \bfL^{p}(\Omega) \right\}.
\end{align}
For a vector field $\bff$ that is a distributional gradient of an object from $W^{1,p}(\Omega)$, $\bff \in \nabla W^{1,p}(\Omega)$, we are \cye{interested in} \emph{potentials}, i.e., scalar-valued functions $\Phi_{\grad}( \bff ) \in W^{1,p}(\Omega)$ such that $\nabla \Phi_{\grad}( \bff ) = \bff$. Then, 
%
\begin{align*}
    \Phi_{\grad} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)
\end{align*}
is the \emph{potential operator}. A particular example of our interest is
%
\begin{align}\label{math:pot_grad_def}
    \Phi_{\grad}( \bff ) := \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } \| u \|_{L^{p}(\Omega)}.
\end{align}
%
Actually, in the present gradient setting and since $\Omega$ is connected, already the gradient constraint determines $\Phi_{\grad}( \bff )$ up to a constant, so that~\eqref{math:pot_grad_def} is a one-dimensional minimization problem.

We want to see whether the $L^{p}(\Omega)$ norm of the potential $\Phi_{\grad}( \bff )$ is bounded in terms of the $\bfL^{p}(\Omega)$ norm of the source field $\bff$, i.e., we want to evaluate the \emph{operator norm}
\begin{align}\label{math:pot_grad_norm}
    C_{\grad,\Omega,p} := \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} }.
\end{align}
This operator norm is indeed finite and it equivalently turns out to be the best constant in the \emph{Poincar\'e--Friedrichs inequality}: for every vector field $\bff \in \nabla W^{1,p}(\Omega)$, there exists $u \in W^{1,p}(\Omega)$ such that $\nabla u = \bff$ and 
\begin{align} \label{math:PF_gradient_1}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} 
    \| \bff \|_{\fooL^{p}(\Omega)}.
\end{align}
Actually, \eqref{math:PF_gradient_1} is further equivalent to the more conventional writing
\begin{align} \label{math:PF_gradient_2}
    \min_{ c \in \bbR } \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p}
    \| \nabla u \|_{\fooL^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    .
\end{align}
We call $C_{\grad,\Omega,p}$ the \emph{Poincar\'e--Friedrichs constant} for the gradient operator with exponent $p$. \cye{A concise writing of all~\eqref{math:pot_grad}--\eqref{math:PF_gradient} is
\begin{align} \label{math:pot_PF_grad}
    \min\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \nabla u = \bff } } 
    \| u \|_{L^{p}(\Omega)}
    &\leq 
    C_{\grad,\Omega,p}
    \| \bff \|_{\fooL^{p}(\Omega)}.
\end{align}
for all $\bff \in \nabla W^{1,p}(\Omega)$.}

\cye{We present a more profound discussion below in Section~\ref{section:poincare}, where we also discuss the difference with the Poincar\'e inequality, where the mean value of $u$ is fixed to zero, and address why we later employ linear potential operators to estimate the Poincar\'e--Friedrichs constant and not directly the minimizing but nonlinear (for $p \neq 2$) potential operators such as~\eqref{math:pot_grad_def}.}

\subsection{Conceptual overview: Potentials for the curl and divergence operators and the corresponding Poincar\'e--Friedrichs inequalities} \label{section:intro_curl_div}

Consider the vector-valued spaces
\begin{align}\label{equation:W_curl_div_p} 
    \bfW^{p}(\curl,\Omega) &:= \left\{ \bfu \in L^{p}(\Omega)^{3} \cye{\suchthat} \curl \bfu \in L^{p}(\Omega)^{3} \right\}
    ,
    \\
    \bfW^{p}(\divergence,\Omega) &:= \left\{ \bfu \in L^{p}(\Omega)^{3} \cye{\suchthat} \divergence \bfu \in L^{p}(\Omega) \right\}
    .
\end{align}
The members of these spaces are those vector fields whose distributional curls and divergences are \cye{respectively} in \cye{the} Lebesgue spaces \cye{$\bfL^{p}$ or $L^{p}(\Omega)$}.  
This only requires that certain sums of distributional partial derivatives are integrable, 
and hence these spaces are not classical Sobolev spaces of vector  fields.

\cye{Let a vector field $\bff$ that is a distributional curl of an object from $\bfW^{p}(\curl,\Omega)$, $\bff \in \curl \bfW^{p}(\curl,\Omega)$, and scalar-valued field $f$ that is a distributional divergence of an object from $\bfW^{p}(\divergence,\Omega)$, $f \in \divergence \bfW^{p}(\divergence,\Omega)$, be given.} 
We are interested in constants for inequalities such as 
\begin{align}\label{math:pot_PF_curl_div}
    \min\limits_{ \substack{ \bfu \in \bfW^{p}(\curl,\Omega) \\ \curl \bfu = \bff } } 
    \| \bfu \|_{L^{p}(\Omega)}
    &\leq 
    C_{\curl,\Omega,p}
    \| \bff \|_{L^{p}(\Omega)} \label{math:pot_PF_curl}
    ,
    \\
    \min\limits_{ \substack{ \bfu \in \bfW^{p}(\divergence,\Omega) \\ \divergence \bfu = f } } 
    \| \bfu \|_{L^{p}(\Omega)}
    &\leq 
    C_{\divergence,\Omega,p}
    \| f \|_{L^{p}(\Omega)} \label{math:pot_PF_div}
    ,
\end{align}
and how these constants can be estimated by the operator norms of \emph{linear} potential operators 
\begin{align*}
    &\Phi_{\curl} : \curl \bfW^{p}(\curl,\Omega) \rightarrow \bfW^{p}(\curl,\Omega)
    ,
    \\
    &\Phi_{\divergence} : \divergence \bfW^{p}(\divergence,\Omega) \rightarrow \bfW^{p}(\divergence,\Omega)
    .
\end{align*}
The fundamental difference \cye{with respect to Section~\ref{subsection:intro_grad}} is that the curl and divergence have infinite-dimensional kernels. In Section~\ref{subsection:intro_grad}, the kernel of the gradient is the one-dimensional space of constant functions. It is thus trivially complemented for all $p$, with a canonical choice of projection. By contrast, in the Banach space case $p \neq 2$, it is not immediately evident that the kernels of the curl and divergence operators are complemented, and a canonical projection only exists here in the Hilbert setting $p=2$. \cye{Actually,} the existence of a bounded linear potential operator here is a non-trivial fact even in the Hilbert space case $p = 2$.

\cye{Let for a moment the topology of $\Omega$ be trivial, let $\Omega$ be weakly Lipschitz, and consider the Hilbert case $p=2$. Then, \eqref{math:pot_PF_curl} is related to the so-called Poincar\'e--Friedrichs--Weber inequality
\begin{align}
    \label{math:Weber}
    \| \bfu\|_{L^{2}(\Omega)} \leq \tilde C_{\curl,\Omega,p} \| \curl \bfu \|_{L^{2}(\Omega)},
\end{align}
for all $\bfu \in \bfW^{2}(\curl,\Omega) \cap \bfW^{2}(\divergence,\Omega)$ with either the tangential or the normal trace zero on $\partial \Omega$ and $\divergence \bfu = 0$. 
Yet another writing is 
\label{math:PF_orth} \begin{align}
    \| \bfu\|_{L^{2}(\Omega)} & \leq \tilde C_{\curl,\Omega,p} \| \curl \bfu \|_{L^{2}(\Omega)}, \label{math:PF_orth_curl}\\
    \| \bfu\|_{L^{2}(\Omega)} & \leq \tilde C_{\divergence,\Omega,p} \| \divergence \bfu \|_{L^{2}(\Omega)} \label{math:PF_orth_div}
\end{align}
for respectively all functions $\bfu$ from $\bfW^{2}(\curl,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv$ from $\bfW^{2}(\curl,\Omega)$ with $\curl \bfv = 0$ and all functions $\bfu$ from $\bfW^{2}(\divergence,\Omega)$ such that $\int_{\Omega} \bfu(x) \cdot \bfv(x) \;dx = 0$ for all $\bfv$ from $\bfW^{2}(\divergence,\Omega)$ with $\divergence \bfv = 0$. We refer to Equation~(5) in~\cite{Fried_diff_forms_55}, Equation~(2) in Gaffney~\cite{Gaff_Hilbert_harm_55}, Weber~\cite{Web_compact_Maxw_80}, \cite[Lemmas~3.4 and~3.6]{Gir_Rav_NS_86}, \cite[Proposition~7.4]{Fer_Gil_Maxw_BC_97}, and the references therein. Still} in the Hilbert space case $p=2$, the Friedrichs inequality~\cite{burenkov1998sobolev} over the Sobolev space with \cye{homogeneous} Dirichlet boundary conditions \cye{on $\partial \Omega$} implies, via duality~\eqref{math:pot_PF_div}. However, that easy observation seems restricted to $p=2$.

Not much attention\cye{, however,} seems to have been given to the study of \cye{computable} constants in~\eqref{math:pot_PF_curl_div}, even if the domain $\Omega$ is convex. \cye{Important results in the Hilbert case $p=2$ can be found in~\cite{guerini2004eigenvalue, Paul_Vald_PF_grad_curl_div_20} that we discuss now in the broader setting of} the exterior calculus formalism\cye{, but the Banach case $p \neq 2$ seems largely unexplored.}
